### 2013-01-11 DGE_paper.tex
initial commit
added two tables

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Genome-wide transcriptome variation in chicken line 6 and 7 in response to Marek's disease infection}
}
% Insert Author names, affiliations and corresponding author email.
\\
Author1$^{1}$, 
Author2$^{2}$, 
Author3$^{3,\ast}$
\\
\bf{1} Author1 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{2} Author2 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
\\
$\ast$ E-mail: Corresponding author@institute.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Gene models summary}
The number of genes and isoforms is shown in Table \ref{tab:gene_models}.

\subsection*{Differential gene expression}
The number of differentially expressed genes is shown in Table \ref{tab:dge}.

\subsection*{Differential exon usage}

\subsection*{Polymorphisms in differentially expressed genes/exons}

\subsection*{Functional analysis}

\section*{Discussion}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

%\section*{References}
% The bibtex filename
\bibliography{template}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}

\begin{table}[!ht]
\caption{
\bf{Gene models summary}}
\begin{tabular}{ccc}
\hline
Method & Gene & Isoform \\
\hline
Assembly & 25,290 & 54,044 \\
Cufflinks & 21,345 & 36,218 \\
Combined & 24,980 & 46,613 \\
\hline
\end{tabular}
\begin{flushleft}Number of genes and isoforms from gene models
built from \emph{de novo} assembly and Cufflinks.
Combination of both methods decreased the number of genes
and isoforms by merging fragmented transcripts to
form more complete gene models.
\end{flushleft}
\label{tab:gene_models}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Differentially expressed genes}}
\begin{tabular}{cccccc}
\hline
Sample & Up-regulated & Down-regulated & Total & Refseq & Ensembl\\
\hline
Line 6 & 1,043 & 595 & 1,638 & 984 (60.0\%) & 1,252 (76.4\%) \\
Line 7 & 1,976 & 1,124 & 3,100 & 2,114 (68.2\%) & 2,574 (83.0\%) \\
Both & 681 & 236 & 917 & 664 (72.4) & 771 (84.1\%) \\
\hline
\end{tabular}
\begin{flushleft}
FDR $< 0.1$
\end{flushleft}
\label{tab:dge}
\end{table}
\end{document}