### 2014-06-20 paper.tex
added a major updates

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Transcriptome variation in response to Marek's disease virus early infection}
}
% Insert Author names, affiliations and corresponding author email.
\\
Likit Preeyanon$^{1}$
C. Titus Brown$^{1,2}$
Hans H. Cheng$^{3,\ast}$
\\
\bf{1} Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI, USA
\\
\bf{2} Computer Science and Engineering, Michigan State University, East Lansing, MI, USA
\\
\bf{3} USDA, ARS, Avian Disease and Oncology Laboratory, East Lansing, MI, USA
\\
$\ast$ E-mail: Corresponding hans.cheng@ars.usda.gov
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
Marek's disease (MD) is caused by highly oncogenic Marek's disease virus (MDV).
Understanding the genetic basis for non-MHC resistance would be of both fundamental and applied importance.
In this study, using computational approaches of RNA-Seq data, we identified
differentially expressed genes and isoforms in response to MDV infection
from highly inbred chickens lines that are MD resistant and susceptible,
and performed pathway and functional analysis.
Results show that expression of genes and isoforms involved in cell cytoskeleton and
adhesion molecules in line 6 may contribute to resistance of the disease by limiting
T cell activation, which in turn prevents the spread of MDV to activated T cells.
In addition, we identified exonic single nucleotide polymorphisms (SNPs) between
lines 6 and 7 predicted their role in regulation of alternative splicing.
Based on our results, we speculate that alternative splice forms of cell adhesion molecules
and cytoskeleton play and important role in controlling T cell activation in line 6.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}

Marek's disease (MD) is an economically significant chicken disease that affects the
poultry industry worldwide with estimated annual loss of
\$2 billion loss has been reported from outbreaks~\cite{morrow2004marek}.
The disease is caused by the highly oncogenic Marek's disease virus (MDV)
an alphaherpesvirus that induces T-cell lymphomas in susceptible birds.
Vaccination is the primary control measure, which is effective in reducing
incidence of tumor formation.
However, since MD vaccines are not sterilizing, they do not prevent infection or
horizontal spread of the virus.
As a consequence, MDV field strains that overcome vaccinal protection have arisen repeatedly over time.
Therefore, there is a need for sustainable alternative controls measures, such as improving genetic
resistance.

Many studies have reported strong associations between MHC alleles and resistance or susceptibility to
MD. For example, chickens with MHC allele B$^{21}$ are highly resistant in contrast to chickens with
B$^{16}$ allele which are highly susceptible.
ADOL lines 6 and 7, both share (B$^6$) allele, yet exhibit different phenotypic responses to
MD; e.g., challenge with JM/102W strain with result in 0 and 100\% MD incidence for lines 6 and 7,
respectively.
Thus, the major unanswered questions are what genetic factors, especially those that are non-MHC,
contribute to susceptibility and resistance of the disease and what is the main contributing
mechanism.

In the past decades, significant efforts have been spent to study variations in a global gene
expression between resistant and susceptible birds using
microarray and RNA-Seq methods in order to identify non-MHC genes that contribute
to resistance to MD~\cite{}.
However, none of the studies have investigated differential expression of alternative isoforms,
which are known to play a significant role in many biological events including
immune responses.
In addition, studies have shown that isoform expression levels can provide better signatures for some
diseases~\cite{zhang2013isoform}.
Changes of isoform expression levels are governed partly by two types of \textit{cis-}regulatory
elements: Exon Splicing Enhancer (ESE) and Exon Splicing Silencer (ESS),
which are located within an exon sequence.
A number of sequence motifs of ESE and ESS have been identified in human and some other organisms
and could be predicted \textit{in silico}.
Mutations that disrupt or create those motifs could alter splicing
patterns leading to aberrant alternative splicing.
A number of disease-associated single-nucleotide polymorphsims in coding regions
(cSNPs) that affect ESEs and ESSs have been well
characterized~\cite{blencowe2000exonic, wang2007splicing}.
%In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing via this mechanism.
Therefore, variations in isoform expression could lead to identification of SNPs
that underlie genetic resistance to MD.
In this article, we reported differential-expressed genes and isoforms that may contribute to
resistance to MD as well as cSNPs that can potentially affect isoform expression levels.

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Differential gene expression in response to MDV infection}

To extensively study gene and isoform expression, we incorporated \textit{Ensembl} gene models
release 73 with \textit{de novo} and a reference-guided transcriptome assembly to build custom
gene models.
The models, therefore, include both \textit{Ensembl} annotated transcripts
and putative transcripts and genes.
The advantage of using custom gene models is it allows an investigation of unannotated genes and
isoforms, which is necessary for in-depth transcriptome study of an immune system.

Many genes were found to be differentially expressed (DE) between control and infected
chickens in lines 6 and line 7.
While the number of unique downregulated genes in both lines was approximately equal,
the number of unique upregulated genes in line 7 was much greater compared
to line 6 (Figure~\ref{degenes_venn})~\cite{lee1983ontogeny}.
Interestingly, some genes that were differentially expressed in both lines were
regulated in the opposite direction (Table~\ref{tab:opposite}).
Among genes downregulated in line 6 but upregulated in line 7 were
\textit{LL} (lung lectin) and \textit{SFTPA1}, which encode a calcium-dependent C-type
lectin and a lung surfactant protein respectively.
Both molecules are important in innate immunity~\cite{hogenkamp2008chicken,kingma2006defense}.
\textit{LIMS1} is involved in cell differentiation and proliferation and
\textit{PPARG} is a suppressor of \textit{NF$\kappa$B}-mediated proinflamatory response.
On the other hand, nearly all genes upregulated in line 6 but downregulated in line 7
are involved in cell survival such as mRNAs splicing, cell growth, and protein synthesis,
except CD7 whose function is involved in T cell-B cell interaction.
This difference suggests that even at this stage of infection in line 6, the lytic phase 
could be subsided, therefore, only genes important for survival of cell are upregulated in line 6.
In comparison, the lytic phase in line 7 may still continue and as a result, genes involved in
immune responses are still upregulated.
In addition, type I interferon (\textit{IFN-$\gamma$} and \textit{IFN-$\beta$}) as well as \textit{INF-$\alpha$3} were found to be highly
upregulated in infected chickens in both lines 6 and 7 (Table~\ref{tab:cytokines}).
However, expressions of genes encoding their corresponding receptors
were not different in line 6, but upregulated in line 7.
This could also reflect the ongoing immune responses in line 7.

Some DE genes have been reported from previous microarray studies were also found to be
differentially expressed in this study.
For example, B6.1 (Bu-1) gene is known to be down-regulated approximately 2.27 fold in susceptible
chicken with the MHC allele B\textsuperscript{19} at 4 d.p.i~\cite{sarson2008transcriptional}.
It was also found to be down-regulated about 3-fold in line 7 (susceptible).
Similarly, \textit{GMZA} gene reported to be upregulated across genetically different chickens
(B$^{19}$, B$^{21}$ alleles), was also found to be highly upregulated.
In contrast, some genes that have been reported to be highly expressed in resistant chickens were
downregulated in both lines. Those genes are \textit{AMIGO2, MMP13} and \textit{CLEC3B}, which were
found to be downregulated more than 2-fold~\cite{sarson2008transcriptional}.
% Several genes involved in innate immune response were differentially expressed in resistant chickens:
% \textit{DDT, NMU, VIPR2} and \textit{HSP5}. Of all those genes, only HSP5 was upregulated.
Other immune genes reported to be highly expressed in susceptible chickens (line 7$_{2}$) including
\textit{AVD, ART1, NOS2, CXCL13L2, MX1} and \textit{SOCS1}~\cite{smith2011systems}
were also found to be highly upregulated in both line 6 and 7.
However, our results show the similar expression patterns for \textit{IL6} and \textit{IL18}, which were
only upregulated in line 7 in early stage of infection (3-5 d.p.i).

In contrast, \textit{IL15} has been reported non differential-expressed between control
and infected chickens in
line 6$_{1}$ and 7$_{2}$~\cite{kaiser2003differential}; however,
it was only upregulated in line 7 in our study.
Expression of \textit{IL15} is induced by \textit{TLR9}, which binds to non-methylated CpG
residues present in genomes of many DNA viruses, including herpes simplex virus.
This cytokine auto-regulates the expression of \textit{CD40}, which is
a transmembrane receptor required for activation of macrophages by CD4 T cells.
Consequently, \textit{CD40} was only upregulated in line 7 (data not shown).

\subsection*{Functional analysis of differential-expressed genes}

To determine pathways that were perturbed during the infection, data were
analyzed by GOSeq, which accounted for gene lengths bias unique for RNA-Seq data~\cite{young2010method}.
Significantly perturbed pathways (FDR $< 0.1$) from both lines that involved in
immune response include TLR signaling pathway, cytokine-cytokine receptor interaction,
intestinal immune network for IgA production, and cell adhesion molecules (CAMs)~\ref{line67_kegg}.
Some other pathways important in response to viral infection and only significantly
enriched in line 7 include phagosome, apoptosis, RIG-I-like receptor signaling pathway,
NOD-like receptor signaling pathway, and lysosome.
Figure~\ref{kegg_phagosome} is a pictorial example of differentially expressed
genes in the phagosome pathway.
Although MHC I gene (\textit{BF1}) was differentially expressed in both lines, other genes involved in
expressing newly synthesized MHC I were only upregulated in line 7
suggesting that new MHC I molecules were actively produced and expressed on cell surface.
% (the minor class I is upregulated but the major class I is downregulated)
Furthermore, gene ontology analysis of biological processes (GO:BP) shows that categories involved in
both adaptive and innate immune responses were enriched in line 7 (supplementary material).
On the other hand, only categories involved in innate immune responses were enriched in line 6.
In addition, enrichment of apoptosis pathway in line 7 indicates that the programmed
cell death could be induced by CTL response to eliminate the viruses.

At this stage of infection, our results suggest that lytic infection of MDV stimulates
both innate and adaptive immune responses, which leads to activation of T cells in line 7.
Only activated T cells are infected by MDV, therefore, the lytic phase could facilitate the spread of
the viruses by enhancing expansion of activated T cells.
Due to cell-associate nature of MDV, the viruses transfer to T cells via cell-to-cell
contact between B cells and
T cells during antigen presentation or B cells activation by T helper cells.
Therefore, it is beneficial for the host to restrain such contact.
However, it is not clear how chickens in line 6 control the lytic infection of MDV.
% Two mechanisms have been speculated to contribute to MD resistance.
% First innate immune responses could be highly effective and could activate
% strong adaptive immune responses that rapidly control viral replication and force the viruses to undergo latent phase.
% Second, innate immune resposnes itself could be highly effective in limiting viral replication~\cite{smith2011systems}.

\subsection*{Differential exon usage (DEU)}

The immune system is isoform-rich and many genes express different isoforms with
distinctive functions in response to stimuli such as stress, chemicals and infection.
Changes in expression of splice forms of immune related genes have been reported to
be associated with increase susceptibility and poor prognosis of diseases~\cite{lynch2004consequences}.
Studying differential isoform expression could therefore shed lights into inherent differences between 
lines 6 and 7 that confer resistance or susceptibility to MD.

In the past decades, microarray technology has been used to study gene and isoform
expressions in many studies, but its sensitivity for detection of structurally similar
isoforms is low, and known or predicted annotations are required to design probes~\cite{kane2000assessment}.
Although RNA-Seq method can provide a reliable estimate of an exon expression compared
to microarray~\cite{pan2008deep} and is not constrained to those limitations, studying isoform
expressions using RNA-Seq is still not straightforward because of the short read lengths.
Reads from current Illumina technology are generally not long enough to span across all exons in an isoform.
In most cases, only exons in close proximity are covered by the same read, which makes it
difficult to accurately predict a full structure of the isoform.
In addition, some genes are fused due to overlapping untranslated regions (UTRs),
which can also result in erroneous predicted isoform structures.

Due to those issues, it is not feasible to accurately estimate expression of isoforms, especially when
gene annotation is constructed from \textit{de novo} assembly~\cite{trapnell2013differential}.
To avoid these issues, we chose to study exon expression instead of isoform expression.
Using MISO with exon-centric method, only reads spanning across a few exons are used and only exons
involved in a splicing event are examined.
The expression of exon inclusion is calculated as Percent Spliced In (Psi or $\Psi$), which can be used to
infer the portion of transcripts that include the exon in each sample~\cite{Katz:2010iv}.
In this study, we investigated the three most common alternative splicing events in
vertebrates, which are skipped exons (SE), an alternative $3\prime$ (A3SS) and $5\prime$ (A5SS) splice site.

\subsubsection*{DEU genes in response to MDV}

Lists of DEU genes from line 6 that show difference in $\Psi$ greater than 0.20 when compared to line 7
in infected chickens are shown in
Tables~\ref{tab:line67i_diff_line67u_one},~\ref{tab:line67i_diff_line67u_two}
and~\ref{tab:line67i_diff_line67u_three}.
Genes can be categorized roughly into four groups based on the pattern of $\Psi$ across control and
infected birds in both lines.

Group I (Table~\ref{tab:line67i_diff_line67u_one}) includes genes with $\Psi$s that were
up- or down-regulated in infected chickens in line 6 only.
This group includes \textit{BCL11B} (B-cell CLL/lymphoma 11B zinc finger proteins), a B-cell lymphoma
associated C2H2-type zinc finger protein encoding gene, which functions as a tumour-suppressor
in T-cell lymphoma in human.
According to homologous alignments on UCSC genome browser, a splice form with the skipped exon
is similar to mouse \textit{BCL11B isoform b}.
The skipped exon was expressed 30\% in infected line 6 chickens; whereas it was rarely expressed
(4-7\%) in the control line 6 and both groups in line 7.
The skipped exon was not found to encode any known protein domain, however, it is in the middle of
two adjacent C2H2-type finger protein domains.
Another gene related to B and T cell lymphomas is \textit{SIK2} (salt-inducible kinase 2).
This gene has been reported to have a negative effect on T cell lymphomas by limiting the transcription of
HTLV-1 virus~\cite{tang2013lkb1}.
Some other genes are important in pre-mRNA splicing including \textit{TRA2A},
\textit{SRSF6} and \textit{GEMIN6}.
\textit{TRA2A} (transformer-2 protein homolog alpha) encodes RNA recognition motif (RRM),
the skipped exon is not found to encode any known protein domain.
\textit{SRSF6} (SR splicing factor 6) encodes a nuclear protein that belongs to the splicing
factor protein family.
\textit{GEMIN6} plays a role in the assembly of spliceosomal snRNP in cytoplasm.
These genes could play a significant role in regulating inclusion of alternative exons.
A gene in this group that might be important for innate immune responses is \textit{RAC3} (Ras-related C3
botulinum toxin substrate 3).
This gene encodes small GTPases, belonging to Ras family, that regulates a wide variety of cellular events
including cell growth, cytoskeletal reorganization, and the activation of protein kinases.
The role of small GTPases in immune responses is discussed further below.

In group II (Table~\ref{tab:line67i_diff_line67u_two}),
 $\Psi$ values were relatively stable in control and infected chickens within line, but not between lines.
Genes that could play an important role in immune responses are \textit{ITGB2} and \textit{HCK}.
\textit{HCK} transmits signal from cell surface receptors such as \textit{FCGR1A, FCGR2A, IL2, IL6,
IL18}, and integrins (\textit{ITGB1, ITGB2}).
\textit{ITGB2} (CD18) encodes subunit $\beta_{2}$ integrin of \textit{LFA-1} and \textit{CR3} receptors.
\textit{LFA-1} plays an important role in adhesion of lymphocytes with other cells.
\textit{CR3} binds to a vast array of ligands and molecules including complement C3bi, microbial proteins,
ICAM-1 and -2, ECM proteins, and coagulation proteins~\cite{}.
It plays a significant role in neutrophils and monocytes activation including phagocytosis, adhesion and
migration~\cite{}.
The role of \textit{ITGB2} in immune responses is discussed further in the next section.
\textit{DYNLT1,DYNLL2, SEPT11, PFN2}, and \textit{ZDHHC7} are also involved in cell
rearrangement and cytokinesis.
In particular, \textit{DYNLT1} and \textit{DYNLL2} are dynein proteins that have been demonstrated to
regulate T cell activation by driving T cell receptor microclusters (TCR-MCs)
toward the center of immune synapse~\cite{hashimoto2011dynein}.

Group III (Table~\ref{tab:line67i_diff_line67u_three})
includes genes that exhibit differential isoform expression only in infected line 7.
A number of genes in this group encode proteins that are parts of
spliceosome: \textit{SRSF3}, \textit{HNRNPDL},
\textit{SFSWAP}, \textit{THOC1}, \textit{RNPC3} and \textit{SRSF5}.
Two genes involved in cell-cell contact regulated by integrins are also in this group.
\textit{PPP1R12A} is a myosin phosphatase that regulates the interaction of actin
and myosin downstream of GTPase Rho proteins.
\textit{PODXL} encodes PODX-like protein that functions in integrin-dependent
manner as both pro-adhesive and anti-adhesive molecules.
This protein is involved in cell-to-cell contact, cell trafficking, and cancer progression~\cite{}.

The last group (Group IV, Table~\ref{tab:line67i_diff_line67u_three})
only has one gene, \textit{GOSR1}.
The $\Psi$ value of this gene was less than 0.20 cutoff in control and infected chickens in line 6, but it is
significantly different between infected chickens in line 6 and 7.
Also, there is a significant difference between control and infected chickens in line 7.
This gene encodes a trafficking membrane protein important for transporting proteins
from \textit{cis-} to \textit{trans-}golgi network.

\subsubsection*{Roles of \textit{LFA-1} and actin cytoskeleton pathway in T cells activation}

Integrins are a special kind of receptors that transmit signals bidirectionally across the cell membrane.
They are heterodimeric composed of an $\alpha$ (large) and a $\beta$ (small)
subunits~\cite{wang2010immunopathologies}.
$\beta_{2}$ (CD18) subunit encoded by \textit{ITGB2} is expressed on lymphocytes and antigen presenting cells
(APCs) as a component of \textit{LFA-1} and \textit{CR3} receptors.
LFA-1 binds to its ligand ICAM-1 to help form a synapse that bring APC and T cell together to
initiate antigen presentation leading to T cell activation~\cite{dustin2000immunological}.
Absence of LFA-1 leads to impaired functions of lymphocytes in proliferation and
tumor rejection~\cite{scharffetter1998spontaneous,schmits1996lfa}.
Mutations in \textit{ITGB2} gene has been associated with type 1 leukocyte adhesion deficiency (LAD-1),
an autosomal-recessive inherited disease found in few families.
The disease is characterized by impair of lymphocytes in adherent-dependent functions, lack of accumulation
to the site of infection and recurrent bacterial and fungal infection~\cite{springer1987lymphocyte}.

Actin cytoskeleton is very important in T cell activation because it enhances activity of
LFA-1 by increasing its avidity and recruiting signaling molecules necessary for
downstream signaling~\cite{dustin2000immunological, van2000avidity}.
Cytoskeleton proteins binding to cytoplasmic domain of LFA-1 are thought to play an important role in
driving LFA-1 to aggregation on the cell surface, resulting in increased avidity.
Aggregation of LFA-1 has been demonstrated to be essential for lymphocytes to bind to the
ligand~\cite{van1994extracellular}.
Interestingly, \textit{RAC3}, \textit{PFN2}, and \textit{PPP1R12A}, which are involved in
actin cytoskeleton pathway (Figure~\ref{kegg_actin}), also expressed different ratios of alternative splice
forms between lines 6 and 7.
Some of these genes are also co-present in three other pathways
that involved in immune responses (Table~\ref{tab:integrin}).
This could be speculated that pre-mRNA splicing of these genes is co-regulated by splicing
regulators or some genetic factors.

\subsubsection*{Protein domains}
To understand the role of the alternative splice forms on the function of these genes,
transcripts were translated to protein sequences by ESTscan~\cite{iseli1999estscan}.
Protein sequences were then searched for annotated protein domains using
the standalone version of InterPro Scan~\cite{quevillon2005interproscan}.
Besides \textit{ITGB2}, other genes have alternative exons located in coding regions and
could potentially affect functional protein domains in some ways.
The exon with alternative 3$\prime$ splice site of \textit{RAC3} encode part of a protein
domain identified as small GTPase of Ras subfamily (ProSiteProfiles:PS5142 and SMART:SM00173).
Rac3 is highly homologous to Rac1 and has been reported to possess the ability of
promoting membrane ruffling, transformation, activation of
c-Jun transcriptional activity and a co-activator of NF$\kappa$B~\cite{,werbajh2000rac}.
Activated Rac also regulates production of superoxide in neutrophils and macrophages.

The alternative exon of \textit{PFN2} seems to disrupt the coding sequence that
encode profilin domain (Pfam:PF00235).
The profilin domain is essential for almost all organisms and its functions include
regulating actin polymerization, controlling complex network of molecular interaction
and transmitting signals from small-GTPase pathway.
It also binds to Rac effector molecules and a number of other ligands~\cite{witke2004role}.
The skipped exon of \textit{PPP1R12A} encodes part of a protein domain annotated as protein
phosphatase 1 (PIRSF:PIRSF038141).
The functional effects of alternative splicing on these functional domains have to be
further investigated by simulation or an experiment.

% Based on our gene model, the skipped exon of \textit{ITGB2} is located in the 5$\prime$ UTR;
% whereas, the skipped exon of \textit{HCK}, encodes protein tyrosine kinase (Pfam:PF07714).

\subsubsection*{Prediction of \textit{cis}-regulatory elements in alternative splicing exons}

Although alternative splicing of genes in all groups could be regulated by
\textit{cis-} and \textit{trans-}splicing factors,
genes in group II are more likely to be regulated by \textit{cis-}regulatory factors.
The ratios of isoform expression in this group were relatively stable within line,
but were significantly different between lines.
Investigation of nucleotide differences within exons of line 6 and 7 could reveal
a possible role of cSNPs in regulating alternative splicing in this group.
We obtained a sequence of an alternative exon from line 6 and used Human Splicing Finder (HSF)
to determine whether SNPs from line 7 could alter predicted ESEs or ESSs.
Results from some genes involved in cytokinesis are discussed in this section.
Exonic SNPs from line 6 and 7 are listed in Table~\ref{tab:deu_snps}.

For \textit{ITGB2}, a SNP (T) at position 26 of the cDNA from line 6, which corresponds to
position 7,183,696 on chromosome 7 is located in a predicted binding site for
SC35, which is an exon enhancer.
Although the exon is not expressed in line 7, we found that there is no polymorphisms between
lines 6 and 7 according to SNPs data from the genome resequencing.
Therefore, this SNP may not be accounted for exclusion of the exon in
line 7 (Figure~\ref{itgb2}).

Exon sequences of \textit{PFN2} from line 6 and 7 differ at position 2,072 on
cDNA or 23,221,934 on chromosome 9.
A small insertion of two AA nucleotides is predicted to create a new binding site
for Tra2-$\beta$ splicing regulator, which serves as a stabilizer of an enhancer
complex~\cite{lopez1998alternative}.
From exon expression data, $\Psi$ of an exon with alternative splicing increases
from 0.20-0.30 to about 0.50 in line 7.
Tra2 could possibly increase inclusion of the exon with alternative 3$\prime$ splice site via
ESE-dependent 3$\prime$ splice site activation (Figure~\ref{pfn2}).
Even though the linear distance between Tra2-$\beta$ binding site and the alternative
3$\prime$ splice site is greater than 1kb, Tra2 could possibly get close to
the 3$\prime$ splice site in the secondary structure of the mRNA.

Replacement of A with G nucleotide in the skipped exon of \textit{DYNLL2} from line 6
is predicted to slightly alter the binding site of several ESEs as well as to create
a new binding site for 9G8.
This exon is upregulated in line 7 compared to line 6, therefore, the present
of the new binding site for 9G8 exon enhancer helps support the expression results.
In addition, A nucleotide is this position matches the reference
nucleotide, therefore, we could expect this exon to be expressed in other datasets.
According to EST tags on the UCSC genome browser, the exon has been found and sequenced
from chicken eyes (15d post-hatched, EST sequence:DR424100).

For \textit{DYNLT1}, the skipped exon was absent in line 7, therefore, we could not
determine the SNPs from RNA-Seq data.
However, data from DNA resequencing showed that there was a polymorphism at position
51,357,865 on chromosome 3.
Replacing G with C from line 7 is predicted to slightly alter predicted binding sites
for 9G8 as well as SRp55 exon enhancer.
However, whether the alteration down regulates the exon in line 7 is unclear.

There are too many SNPs in the exon of \textit{SEPT11} and \textit{ZDHHC7}, making it
unfeasible to predict which SNP regulates the exon expression.
Therefore, we do not discuss these two genes in this section.
However, results from HSF analysis of these two genes and
other genes in this group are provided in the supplementary materials.

\section*{Discussion}

Custom gene models built from combination of gene models from \textit{de novo} assembly,
reference-based assembly, and Ensembl allowed us to identify genes and isoforms
that might play an important role in resistance to MD.
Results from gene expression analysis indicate that adaptive immune responses were active
during lytic infection in line 7.
Because only activated T cells are infected, elicitation of adaptive immune responses could
help spread the viruses by recruiting and activating more T cells.
Therefore, active immune responses are favorable for the virus.
On the other hand, the delay of adaptive immune responses in line 6, in this case,
could benefit the host by limiting infection of activated T cells.

Using exon-centric approach, we identified several genes involved in
actin cytoskeleton, especially \textit{ITGB2},
that expressed different ratios of alternative isoforms between lines 6 and 7.
As described above, \textit{ITGB2} encodes $\beta$-subunit of integrins including LFA-1, which exclusively
expressed in lymphocytes and plays a major role in lymphoproliferation,
antigen presentation, T cells activation, and cytotoxicity.
In patients with LFA-1 deficiency (LAD), the response of lymphocytes to mitogens is
decreased~\cite{springer1987lymphocyte}.
The decrease in responsiveness to mitogens have been shown to correlate with resistance to MD
by Lee and Bacon~\cite{lee1983ontogeny}, who illustrated that resistant birds
(line 6 and N) were less responsive to phytohemagglutinin (PHA) than susceptible birds (line 7 and P).
Even though the exact mechanism is not known, incompetency of T cells in response to
stimuli appears to benefit resistant birds because MD virus could not induce T cells
to proliferate and cause them to undergo neoplastic transformation.
It has also been suggested that the mechanism that controls
both lymphocyte proliferation induced by MD virus and lymphocyte proliferation induced by
immune response is the same~\cite{pazderka1975histocompatibility}.
Therefore, it may be useful to consider a link between deficiency of lymphocytes in line 6 to
the alternative splice form of \textit{ITGB2} that is only expressed in line 6.
Although the exon included in the alternative splice form is non-coding, it could serve important functions in
translation or posttranscriptional regulation.

Interestingly, by grouping genes based on patterns of $\Psi$s, we found that many genes in group
II: \textit{ITGB2, PFN2, DYNLL2, DYNLT1, SEPT11}, and \textit{ZDHHC7}, are involved in cytokinesis or
cell synapse, which are important for aggregation of LFA-1 and T cell activation as discussed above.
The $\Psi$ values in group II are more similar within line than between lines in both control and
infected groups, suggesting the possibility that the alternative splicing is
controlled by a genetic mechanism.
Based on the independent criteria, enrichment of cytoskeletal genes in this group could not be
just a mere coincidence, but an evidence suggesting that inefficiency of T cell activation
in line 6 is linked to a genetic factor that regulate alternative splicing of \textit{ITGB2}
and other genes in the cytoskeleton system.
Experimental validation of exonic SNPs provided by this study could shed some lights on
underlying polymorphisms that contribute to MD resistance.

To conclude, we attempted to employ custom gene models built from RNA-Seq data to
identify underlying genetic factors contributing to MD resistance.
We identified a number of genes that expressed different ratios of alternative splice forms
between lines 6 and 7.
Interestingly, several genes are involved in actin cytoskeleton and cytokinesis, which have
not been of great interest in the filed of MD research.
Even though we mainly discussed the possible role of \textit{ITGB2} in MD resistance, other
genes cannot be precluded and should be a candidate for further investigation and experimental studies.
Moreover, a full mechanism of MD resistance may be highly complex and require more data from different
stages of infection as well as a greater sequencing depth in order to identify all
genes and isoforms involved.
To enable the study of unannotated gene and isoform expression,
our approach of constructing gene models from RNA-Seq should be iteratively used to extend
the Ensembl gene models.

% However, to mechanism that contribute to resistance to MD could be very complex and integration
% of global gene expression data as well as experimental data from different stages of infection will be
% required to fully elucidate.
% Moreover, better gene models and a reference genome will be necessary.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection{Gene Models Construction}

Due to lack of complete gene models for chickens, we employed two methods to
construct gene models from RNA-Seq reads.
First, short reads were quality trimmed with conditri/2.1~\cite{}
and assembled using Velvet/1.21~\cite{} and Oases/0.2.06~\cite{} to obtain long transcripts.
Assembly was done with hash lengths range from 21 to 31 for both local and global assembly
(described in Gimme paper).
Poly-A tails were trimmed and low complexity transcripts were removed by Seqclean~\cite{}.
All transcripts were then aligned to chicken reference genome (galGal4, with unplaced and random 
chromosomes removed) with BLAT~\cite{}.
Second, reads were aligned to reference genome using Tophat/2.0.5~\cite{} and \textit{Ensemble} gene
models release 73 was used to guide reference-based assembly by Cufflinks2~\cite{}.
Alignments from BLAT and models from Cufflinks were then combined to construct gene models by Gimme
(manuscript in preparation).

\subsection{Differential Gene Expression and Gene Ontology}

To identify DE genes, reads were mapped to transcripts by RSEM v.1.10.1~\cite{li2011rsem}, which is
also used to estimate genes expression and identify DE genes.
Data from single- and paired-end datasets from the same line were treated as biological replicates.
To identify enriched pathways and ontology terms, a list of DE genes was analysed by GOSeq v.1.10.0 based on
chicken KEGG annotations.
P-values were corrected by Benjamini-Hochberg multiple testing correction.
Genes, patways and GO terms with corrected P-value $<0.1$ were considered significant.
Pathview~\cite{luo2013pathview} was used to create a KEGG pathway diagram with colors representing relative level of gene expressions.

\subsection{Differential Splicing Event}
Gene models were converted to alternative splicing models using a Python script.
In order to increase sensitivity, read counts from single- and paired-end samples were combined and treated
as single-end reads for splicing event analysis with MISO/0.4.9~\cite{Katz:2010iv}.
Splicing events with Bayes factor $>10$ and $\Delta\Psi>0.20$ were considered significant.
Read coverages and $\Psi$ distributions were plotted using Sashimi plot~\cite{Katz:2013vx}.

\subsection{Variant Calling and \textit{In Silico} Splicing Analysis}
Variants were called using mpileup command from SAMTools/0.1.18~\cite{} and BCFTools~\cite{}.
Only variants with quality score $\ge20$ were used for mutation analyses.
Exon enhancers and suppressors were predicted using Human Splicing Finder web portal~\cite{}.
Human default parameter settings were used in all analyses.
The following regulatory sequences were used to determined the effect of variants:
HSF integrated matrices for serine/arginine-rich proteins (SRp40, SC35, SF2/ASF, SF2/ASF,
IgM/BRCA1, and SRp55), exonic splicing enhancer (ESE), RESCUE-ESE hexamer (RESCUE-ESE),
putative 8-mer ESEs (PESEs) and putative 8-mers exonic splicing silencers (PESSs),
exon-identity and intron-identiry elements (EIEs and IIEs), hetronuclear ribonucleoprotein-binding
motifs, and Fas exonic splicing silencers.

\subsection{Protein domains search}

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

% \section*{References}
% The bibtex filename
\bibliography{ref}{}

\section*{Figure Legends}

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[width=6in]{degenes_venn.pdf}
    \end{center}
    \caption{
        {\bf Differential-expressed genes in response to MDV infection}
    }
    \label{degenes_venn}
\end{figure}

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[width=7in]{line67_KEGG_cleveland.pdf}
    \end{center}
    \caption{
        {\bf Enriched KEGG pathways.}
    }
    \label{line67_kegg}
\end{figure}

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[width=6in]{gga04145_degenes_multi.png}
    \end{center}
    \caption{
        {\bf Phagosome pathway.}
    }
    \label{kegg_phagosome}
\end{figure}

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[width=6in]{hsa04810_deu_genes.png}
    \end{center}
    \caption{
        {\bf Human regulation of actin cytosekeleton pathway.}
    }
    \label{kegg_actin}
\end{figure}

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[width=6in]{itgb2_miso.pdf}
    \end{center}
    \caption{
        {\bf ITGB2 exon expression.}
        A single nuclotide substition (T$>$C) at position 7,183,696 is predicted to disrupt a putative ESE
        motif which binds SC35 on the skipped exon of IRGB2 gene shown in this figure.
        The prediction correlate with exclusion of the exon in line 7 ($\Psi \leq 0.02$).
        A new binding site for a silencer is also predicted and could also promote the exclusion of the exon. 
    }
    \label{itgb2}
\end{figure}

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[width=6in]{pfn2_miso.pdf}
    \end{center}
    \caption{
        {\bf PFN2 exon expression.}
        A small insertion of AA nucleotides at position 23,221,934 is predicted to create a putative ESE motif
        which binds to Tra2 protein.
        Tra2 could promote inclusion of exon with alternative 3$\prime$ splice site via ESE-dependent
        3$\prime$ splice site activation resulting in higher expression of the exon in line 7.
        Even though the predicted ESE motif is located at a substantial distance from the alternative 3$\prime$ splice site,
        Tra2 could still regulate the splicing if a secondary structure of the mRNA moves the ESE motif closer to the
        alternative 3$\prime$ splice site.
    }
    \label{pfn2}
\end{figure}

\section*{Tables}

% \begin{table}[!ht]
%     \caption{
%     \bf{Gene models summary}}
%     \begin{tabular}{ccc}
%         \hline
%         Method & Gene & Isoform \\
%         \hline
%         Assembly & 25,290 & 54,044 \\
%         Cufflinks & 21,345 & 36,218 \\
%         Combined & 24,980 & 46,613 \\
%         \hline
%     \end{tabular}
%     \begin{flushleft}
%         Number of genes and isoforms from gene models
%         built from \textit{de novo} assembly and Cufflinks.
%         Combination of both methods decreased the number of genes
%         and isoforms by merging fragmented transcripts to
%         form more complete gene models.
%     \end{flushleft}
%     \label{tab:gene_models}
% \end{table}

\begin{table}[!ht]
\caption{
\bf{Genes regulated in opposite directions}}
    \begin{tabular}{cccc}
        \hline
        Gene & Description & $log_{2}$FC & \\
         & & Line 6 & Line 7 \\
        \hline
        LL & Lung lectin & -3.36 & 8.70 \\
        GIF & Gastric intrinsic factor & -2.15 & 3.11 \\
        SFTPA1 & Surfactant protein A1 & -4.84 & 3.73 \\
        SCAF8 & SR-related CTD-associated factor 8 & -8.70 & 8.24 \\
        LIMS1 & LIM and senescent cell antigen-like domains 1 & -2.26 & 1.33 \\
        PPARG & Peroxisome proliferator-activated receptor gamma & -6.99 & 2.06 \\
        C14ORF1 & Chromosome 14 open reading frame 1 & -3.11 & 2.71 \\
        Unknown & Unknown & -3.08 & 1.62 \\
        \hline
        ATP8A2 & ATPase, aminophospholipid transporter, class I, type 8A, member 2 & 7.66 & -7.80 \\
        S1PR1 & Sphingosine-1-phosphate receptor 1 & 1.31 & -6.95 \\
        MED9 & Mediator complex subunit 9 & 8.22 & -3.15 \\
        DNAJA2 & DnaJ (Hsp40) homolog, subfamily A, member 1 & 2.34 & -6.02 \\
        RAD17 & RAD17 homolog (S. pombe) & 1.45 & -1.80 \\
        PSMG3 & Proteosome assembly chaperone 3 & 1.52 & -1.26 \\
        PNISR & PNN-interacting serine/argining-rich protein & 5.58 & -1.81 \\
        THOC7 & THO complex 7 homolog (Drosophila) & 7.00 & -6.90 \\
        YTHDC2 & YTH domain containing 2 & 1.00 & -1.74 \\
        RPL39 & Ribosomal protein L39 & 3.31 & -1.59 \\
        CD7 & CD7 molecule & 2.88 & -1.38 \\
        NDUFB3 & NADH dehydrogenase (uqiquinone) 1 beta subcomplex 3 & 1.03 & -1.34 \\
        LOC100858785 & Unknown & 1.26 & -1.75 \\
        Unknown & Unknown & 6.30 & -5.22 \\
        \hline
    \end{tabular}
    \begin{flushleft}
        (-) down-regulated, (+) up-regulated
    \end{flushleft}
    \label{tab:opposite}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Cytokine-related gene expression in response to MDV infection}}
    \begin{tabular}{cccccc}
        \hline
        & & $log_{2}$FC & \\
        Symbol & Description & Line 6 & Line 7 \\
        \hline
        IL2RG & Interleukin 2 receptor, $\gamma$ & -- & 0.55 \\
        IL6 & Interleukin 6 (interferon, $\beta$ 2) & -- & 5.15 \\
        IL6ST & Interleukin 6 signal transducer (gp130, oncostatin M receptor) & -- & 1.36 \\
        IL8L1 & Interleukin 8-like 1 & 1.90 & -- \\
        IL15 & Interleukin 15 & -- & 1.06 \\
        IL18 & Interleukin 18 (interferon-$\gamma$ inducing factor) & 1.92 & 4.06 \\
        IL18R1 & Interleukin 18 receptor 1 & 1.94 & 1.64 \\
        IFNG & Interferon-$\gamma$ & 5.14 & 4.90 \\
        IFNB & Interferon-$\beta$ & 4.83 & 5.64 \\
        IFNA3 & Interferon-$\alpha$ 3 & 4.09 & 5.48 \\
        IFNGR1 & Interferon-$\gamma$ receptor 1 & -- & 2.05 \\
        IFNGR2 & Interferon-$\gamma$ receptor 2& -- & 0.50 \\
        IFNAR1 & Interferon-$\alpha$,$\beta$ receptor 1 & -- & 1.46 \\
        IFNAR2 & Interferon-$\alpha$,$\beta$ receptor 2 & -- & 0.58 \\
        \hline
    \end{tabular}
    \begin{flushleft}
    \end{flushleft}
    \label{tab:cytokines}
\end{table}

\begin{table}[!ht]
\caption{
\bf{DEU between Line 6 and Line 7 in infected birds group I}}
\begin{tabular}{cccccccc}
\hline
& & & & Line 6 ($\Psi$) & & Line 7 ($\Psi$) & \\
Type & Event ID & Ensembl & Symbol  & Un & Inf & Un & Inf \\
\hline
SE & chr2:13729.ev2 & ENSGALG00000010973 & \textbf{TRA2A} & 0.45 & \textbf{0.75} & 0.58 & 0.50 \\
SE & chr5:22858.ev1 & ENSGALG00000011127 & BCL11B & 0.07 & \textbf{0.30} & 0.06 & 0.04 \\
SE & chr2:13065.ev1 & ENSGALG00000013137 & INO80C & 0.14 & \textbf{0.35} & 0.96 & 0.86 \\
SE & chr20:14995.ev1 & ENSG00000124193* & SRSF6 & 0.43 & \textbf{0.72} & 0.54 & 0.34 \\
A5SS & chr7:24049.ev1 & ENSGALG00000009824 & \textbf{C7H2ORF77} & 0.50 & \textbf{0.73} & 0.32 & 0.38 \\
A5SS & chr3:18098.ev1 & ENSGALG00000013821 & \textbf{GEMIN6} & 0.84 & \textbf{0.61} & 0.81 & 0.85 \\
A3SS & chr1:6450.ev1 & ENSGALG00000027665 & \textbf{SYNGR1} & 0.46 & \textbf{0.22} & 0.68 & 0.60 \\
A3SS & chr2:13270.ev1 & ENSGALG00000026498 & \textbf{Unknown} & 0.12 & \textbf{0.71} & 0.10 & 0.34 \\
A3SS & chr19:12399.ev1 & ENSGALG00000005685 & \textbf{KSR1} & 0.23 & \textbf{0.56} & 0.28 & 0.35 \\
A3SS & chr1:6437.ev1 & ENSGALG00000012050 & \textbf{TNRC6B} & 0.57 & \textbf{0.39} & 0.95 & 0.93 \\
A3SS & chr18:11800.ev1 & ENSGALG00000002859 & \textbf{RAC3} & 0.31 & \textbf{0.15} & 0.33 & 0.39 \\
\hline
\end{tabular}
\begin{flushleft}
    *human homologs, Un=uninfected, Inf=infected.
    Bold face indicates that there is a SNP between line 6 and 7 within an alternative exon.
\end{flushleft}
\label{tab:line67i_diff_line67u_one}
\end{table}

\begin{table}[!ht]
\caption{
\bf{DEU between Line 6 and Line 7 in infected birds group II}}
\begin{tabular}{cccccccc}
\hline
& & & & Line 6 ($\Psi$) & & Line 7 ($\Psi$) & \\
Type & Event ID & Ensembl & Symbol  & Un & Inf & Un & Inf \\
\hline
SE & chr17:11370.ev2 & ENSGALG00000004971 & URM1 & \textbf{0.07} & \textbf{0.03} & 0.18 & 0.23 \\
SE & chr2:12495.ev3 & ENSGALG00000005582 & \textbf{KLHL18} & \textbf{0.44} & \textbf{0.27} & 0.56 & 0.52 \\
SE & chr7:24327.ev1 & ENSGALG00000007511 & \textbf{ITGB2} & \textbf{0.17} & \textbf{0.22} & 0.02 & 0.01 \\
SE & chr2:12984.ev1 & ENSGALG00000012809 & ECI2 & \textbf{0.15} & \textbf{0.29} & 0.58 & 0.49 \\
SE & chr8:25247.ev1 & ENSGALG00000028790 & DNASE2B & \textbf{0.38} & \textbf{0.50} & 0.29 & 0.26 \\
SE & chr1:4653.ev1 & ENSGALG00000011682 & CNOT4 & \textbf{0.57} & \textbf{0.63} & 0.40 & 0.40 \\
SE & chr20:14907.ev3 & ENSGALG00000006522 & \textbf{HCK} & \textbf{0.46} & \textbf{0.59} & 0.99 & 0.97 \\
SE & chr4:21539.ev1 & ENSGALG00000015709 & \textbf{TACC3} & \textbf{0.87} & \textbf{0.94} & 0.76 & 0.72 \\
SE & chr2:32201.ev1 & ENSGALG00000012258 & GOLGA4 & \textbf{0.33} & \textbf{0.34} & 0.51 & 0.61 \\
SE & chr3:19470.ev1 & ENSGALG00000019979 & DYNLT1 & \textbf{0.23} & \textbf{0.22} & 0.02 & 0.02 \\
SE & chr19:12391.ev1 & ENSGALG00000005522 & \textbf{DYNLL2} & \textbf{0.01} & \textbf{0.02} & 0.19 & 0.25 \\
A5SS & chr6:23812.ev1 & ENSG00000107651* & \textbf{SEC23IP} & \textbf{0.02} & \textbf{0.12} & 0.50 & 0.43 \\
A5SS & chr2:12775.ev1 & ENSGALG00000011488 & \textbf{CMTM7} & \textbf{0.55} & \textbf{0.67} & 0.37 & 0.41 \\
A3SS & chr8:25257.ev1 & ENSGALG00000008939 & FUBP1 & \textbf{0.58} & \textbf{0.74} & 0.41 & 0.46 \\
A3SS & chr9:25750.ev1 & ENSGALG00000010410 & \textbf{PFN2} & \textbf{0.71} & \textbf{0.78} & 0.53 & 0.50 \\
A3SS & chr4:21230.ev1 & ENSGALG00000011476 & \textbf{SEPT11} & \textbf{0.22} & \textbf{0.14} & 0.40 & 0.42 \\
A3SS & chr4:20185.ev1 & ENSGALG00000008507 & THOC2 & \textbf{0.48} & \textbf{0.47} & 0.31 & 0.22 \\
A3SS & chr11:8703.ev1 & ENSGALG00000020987 & \textbf{ZDHHC7} & \textbf{0.42} & \textbf{0.23} & 0.57 & 0.55 \\
A3SS & chr4:21136.ev1 & ENSGALG00000027908 & \textbf{LOC422528} & \textbf{0.72} & \textbf{0.61} & 0.87 & 0.91 \\
\hline
\end{tabular}
\begin{flushleft}
    *human homologs, Un=uninfected, Inf=infected.
    Bold face indicates that there is a SNP between line 6 and 7 within an alternative exon.
\end{flushleft}
\label{tab:line67i_diff_line67u_two}
\end{table}

\begin{table}[!ht]
\caption{
\bf{DEU between Line 6 and Line 7 in infected birds group III and IV}}
\begin{tabular}{cccccccc}
\hline
& & & & Line 6 ($\Psi$) & & Line 7 ($\Psi$) & \\
Type & Event ID & Ensembl & Symbol  & Un & Inf & Un & Inf \\
\hline
SE & chr12:8987.ev1 & ENSGALG00000008320 & EDEM1 & 0.96 & 0.99 & 0.91 & \textbf{0.72} \\
SE & chr4:20411.ev1 & ENSGALG00000023199 & HNRPDL & 0.39 & 0.40 & 0.30 & \textbf{0.18} \\
SE & chrZ:26257.ev1 & ENSGALG00000001745 & PSTPIP2 & 0.06 & 0.04 & 0.13 & \textbf{0.26} \\
SE & chr1:6316.ev3 & ENSG00000058272* & PPP1R12A & 0.99 & 0.97 & 0.96 & \textbf{0.77} \\
SE & chr1:4478.ev2 & ENSGALG00000009029 & TSPAN12 & 0.09 & 0.15 & 0.20 & \textbf{0.46} \\
SE & chr4:20075.ev1 & ENSGALG00000006157 & DDX26B & 0.67 & 0.61 & 0.58 & \textbf{0.84} \\
SE & chr6:23515.ev1 & ENSGALG00000003861 & HERC4 & 0.31 & 0.37 & 0.45 & \textbf{0.06} \\
SE & chr26:16840.ev1 & ENSGALG00000000533 & SRSF3 & 0.36 & 0.38 & 0.30 & \textbf{0.16} \\
SE & chr11:8507.ev2 & ENSGALG00000000904 & \textbf{C11H16ORF57} & 0.91 & 0.98 & 0.84 & \textbf{0.78} \\
SE & chr1:4323.ev1 & ENSGALG00000006409 & PODXL & 0.21 & 0.34 & 0.26 & \textbf{0.13} \\
SE & chrZ:26582.ev2 & ENSGALG00000014642 & LOC374195 & 0.60 & 0.57 & 0.70 & \textbf{0.81} \\
SE & chr6:23833.ev1 & ENSG00000175029* & CTBP2 & 0.38 & 0.38 & 0.23 & \textbf{0.12} \\
A5SS & chr15:10720.ev1 & ENSGALG00000002487 & \textbf{SFSWAP} & 0.58 & 0.73 & 0.55 & \textbf{0.41} \\
A5SS & chr7:24350.ev1 & ENSGALG00000008038 & \textbf{SF3B1} & 0.41 & 0.57 & 0.55 & \textbf{0.31} \\
A3SS & chr2:13147.ev1 & ENSGALG00000014915 & \textbf{THOC1} & 0.33 & 0.48 & 0.33 & \textbf{0.23} \\
A3SS & chr23:15908.ev1 & ENSGALG00000000720 & LOC419563 & 0.12 & 0.06 & 0.17 & \textbf{0.30} \\
A3SS & chr8:25157.ev1 & ENSGALG00000005162 & RNPC3 & 0.45 & 0.64 & 0.58 & \textbf{0.33} \\
A3SS & chrZ:27197.ev1 & ENSGALG00000000189 & \textbf{YTHDC2} & 0.44 & 0.59 & 0.42 & \textbf{0.32} \\
A3SS & chr23:15983.ev1 & ENSG00000163875* & \textbf{MEAF6} & 0.28 & 0.57 & 0.40 & \textbf{0.29} \\
A3SS & chr5:21970.ev1 & ENSGALG00000009421 & \textbf{SRSF5} & 0.55 & 0.72 & 0.54 & \textbf{0.39} \\
\hline
SE & chr27:17351.ev1 & ENSGALG00000001107 & GOSR2 & 0.73 & 0.92 & 0.38 & 0.60 \\
\hline
\end{tabular}
\begin{flushleft}
    *human homologs, Un=uninfected, Inf=infected.
    Bold face indicates that there is a SNP between line 6 and 7 within an alternative exon.
\end{flushleft}
\label{tab:line67i_diff_line67u_three}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Pathways containing \textit{RAC3, ITGB2, PFN2} and \textit{PPP1R12A}}}
\begin{tabular}{ccc}
\hline
Pathway ID &  Description & Gene \\
\hline
hsa04810 & Regulation of actin cytoskeleton & \textit{RAC3, ITGB2, PPP1R12A, PFN2} \\
hsa04015 & RAP1 signaling pathway & \textit{RAC3, ITGB2, PFN2} \\
hsa04650 & Natural killer cells cytotoxicity & \textit{RAC3, ITGB2} \\
hsa05416 & Viral myocarditis & \textit{RAC3, ITGB2} \\
hsa04510 & Focal adhesion & \textit{RAC3, PPP1R12A} \\
\hline
\end{tabular}
\begin{flushleft}
\end{flushleft}
\label{tab:integrin}
\end{table}

\begin{table}[!ht]
\caption{
\bf{SNPs in exons}}
\begin{tabular}{ccccccc}
\hline
Gene &  Chromosome & Position & Reference & Line 6 & Line 7 & Strand \\
\hline
ITGB2 & 7 & 7183696 & C & \textit{T} & . & - \\
PFN2 & 9 & 23221934 & -  - & . & \textit{AA} & + \\
DYNLT1 & 3 & 51357865 & C & . & \textit{G} & - \\
DYNLL2 & 19 & 8694148 & G & . & \textit{A} & - \\
\hline
\end{tabular}
\begin{flushleft}
\end{flushleft}
\label{tab:deu_snps}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Results from human splicing finder}}
\begin{tabular}{|l|p{1cm}|p{3cm}|l|l|l|l|}
\hline
Gene &  cDNA Position & Linked SR protein & Type & Reference Motif & Mutant Motif & Variation \\
\hline
% ITGB2 & 20 & SC35 & ESE$^{1}$ & TGCTCATG (78.19) & & Site broken \\
%  & 22 & SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & CTCATGG (77.23) & CTCACGG (91.15) & +18.03\% \\
%  & 22 & SF2/ASF(IgM-BRCA1), SF2/ASF & ESE$^{1}$ & CTCATGG (77.23) & CTCACGG (89.23) & +15.53\% \\
%  & 22 & SF2/ASF, SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & CTCATGG (74.55) & CTCACGG (91.15) & +22.27\% \\
%  & 22 & SF2/ASF, SF2/ASF & ESE$^{1}$ & CTCATGG (74.55) & CTCACGG (89.23) & +19.69\% \\
%  & 24 & SRp55 & ESE$^{1}$ & & CACGGA (79.30) & New site \\
%  & 26 & SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & & CGGAGAT (80.00) & New site \\
%  & 26 & SF2/ASF & ESE$^{1}$ & & CGGAGAT (75.36) & New site \\
%  & 21 & & ESS$^{3}$ & & GCTCACGG (63.35) & New site (-5.59) \\
%  & 23 & & ESS$^{3}$ & TCATGGAG (61.41) & & Site broken (3.16) \\
%  & 23 & & ESS$^{3}$ & ATGGAGAT (64.83) & ACGGAGAT & -5.16\% \\
%  & 24 & & ESS$^{4}$ & CATGGA (65.48) & CACGGA (65.48) & 0\% \\
DYNLT1 & 59 & SRp55, SRp40 & ESE$^{1}$& TGAATC (78.47) & TGAATGC (79.10) & +0.81\% \\
& 60 & 9G8 & ESE$^{2}$ & GAATCC (60.07) & GAATGC (63.49) & +5.70\% \\
\hline
DYNLL2 & 2 & SF2/ASF (IgM-BRCA1) & ESE$^{1}$ & CTCCGGG (86.38) & CTCCGAG (72.69) & -15.85\% \\
& 2 & SF2/ASF, SF2/ASF (IgM-BRCA1) & ESE$^{1}$ & CTCCGGG (79.91) & CTCCGAG (72.69) & -9.03\% \\
& 4 & SF2/ASF (IgM-BRCA1) & ESE$^{1}$& CCGGGGT (73.00) & CCGAGGT (86.23) & 18.12\% \\
& 4 & SF2/ASF (IgM-BRCA1), SF2/ASF & ESE$^{1}$ & CCGGGGT (73.00) & CCGAGGT (82.94) & 13.61\% \\
& 6 & 9G8 & ESE$^{2}$ & & GAGGTG (60.67) & New site \\
& 6 & hnRNP A1 & ESS$^{4}$& & GAGGTG (74.05) & New site \\
\hline
PFN2 & 2068 & Tra2-$\beta$ & ESE$^{1}$ & AAAAT (81.02) & AAAAa & +16.19\% \\
& 2069 & Tra2-$\beta$ & ESE$^{1}$ & & AAAaa (94.14) & New site \\
& 2070 & Tra2-$\beta$ & ESE$^{1}$ & & AAAaaT (81.02) & New site \\
 & 2066 & & ESS$^{3}$ & & ACAAAAaa (38.13) & New site \\
 & 2067 & & ESS$^{3}$ & & CAAAAaaT (28.85) & New site \\
\hline

\end{tabular}
\begin{flushleft}
    $^{1}$ESE Finder matrices for SRp40, SC35, SF2/ASF and SRp55 proteins.
    $^{2}$ESE motifs from HSF.
    $^{3}$Predicted PESS Octamers from Zhang \& Chasin.
    $^{4}$hnRNP motif.
\end{flushleft}
\label{tab:spliceosome}
\end{table}

\end{document}